\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}
\usepackage{multirow}

\title{\textbf{nous on MemoryAgentBench:\\A Paper-Faithful Evaluation of a Cognitive Memory Agent}}
\author{Timur Fatykhov\\\small\textit{with Claude (Anthropic) as evaluation engineer}}
\date{July 4, 2026}

\newcommand{\nous}{\textsf{nous}}
\newcommand{\mab}{MemoryAgentBench}

\begin{document}
\maketitle

\begin{abstract}
\noindent We evaluate \nous{}, a cognitive memory agent with episodic ingestion,
LLM-based fact extraction, and sleep-cycle consolidation, on \mab{}
(Hu et al., ICLR 2026), a benchmark that measures four memory competencies of
conversational agents: Accurate Retrieval (AR), Test-Time Learning (TTL),
Long-Range Understanding (LRU), and Conflict Resolution (CR).
We reproduce the benchmark's evaluation protocol exactly---its per-dataset answer
prompts, its per-dataset graders (including two GPT-4o judge protocols), vendored
from the benchmark's open-source repository and verified line-by-line by an
independent review---and drive a real \nous{} HTTP server through its full
production cognitive cycle for every test instance.
On the evaluated subsets, \nous{} scores
\textbf{AR 0.836} (n=568, 95\% CI [0.81, 0.87]),
\textbf{CR 0.725} (n=320, CI [0.68, 0.77]; single-hop 0.887, multi-hop 0.562),
\textbf{LRU 0.824} (n=68, CI [0.73, 0.91]),
and \textbf{TTL 0.555} (n=200, CI [0.49, 0.62]).
The first three are robustly above every method in the benchmark paper's
results (best per-competency averages there: 71.8\%, 29.5\%, 62.2\%); TTL sits
at the top of that range (53.9\%). Against the 2026 field at sub-dataset
level --- with every already-ingested context now at \emph{full} question
coverage, under a reconstruction of its own gpt-5-judge protocol, and with
margins graded against known backbone swings ($\sim$9.7pp) --- \nous{} and
the leading 2026 system are statistical peers overall: one decidable \nous{}
win (multi-hop conflict resolution, $+$21.2pp), one decidable loss
(in-context test-time learning, $-$21.5pp), one weak cell (LongMemEval,
$-$7.6pp), and everything else (event retrieval $+$8.9, single-hop document
QA $+$8.0, detective QA $+$8.1, single-hop conflict $+$8.4, multi-hop
document QA $+$2.0) within backbone-swing range.
We describe the testing technique, a statistical-validation pass that materially
corrected one competency's initial estimate, and the threats to validity that
qualify a direct leaderboard comparison.
\end{abstract}

\section{Introduction}

Long-running LLM agents need memory that outlives a single context window:
they must store what they were told, learn new skills from it, reason over
inputs far larger than the model context, and revise beliefs when information
changes. \mab{}~\cite{mab} operationalizes these requirements as four
competencies and evaluates them by streaming large corpora into an agent
incrementally and then asking questions against what the agent retained.

\nous{} is a cognitive AI agent whose memory subsystem combines an episodic
store (conversation episodes and summaries), a semantic store (LLM-extracted
facts with typed graph edges), full-fidelity content chunks with hybrid
(vector + lexical) retrieval, and an offline \emph{sleep} consolidation cycle
that links, deduplicates, and densifies memory between sessions. This report
documents a from-scratch evaluation harness that measures \nous{} on \mab{}
under its \emph{production} configuration, and places the results against the
methods evaluated by the benchmark's authors.

Two principles governed the effort:
\begin{enumerate}
  \item \textbf{Paper fidelity.} The score must be computed exactly as the
        benchmark computes it---same answer prompts, same graders, same judge
        models---so numbers are comparable. All grading code was vendored from
        the benchmark's MIT-licensed repository and cross-reviewed against it
        by an independent reviewer agent with no stake in the outcome.
  \item \textbf{Statistical honesty.} No above/below-field claim is made unless
        the 95\% confidence interval of the measured accuracy clears the
        best published field value. This rule changed one headline conclusion
        (Section~\ref{sec:validation}).
\end{enumerate}

\section{System Under Test}

\subsection{The \nous{} cognitive cycle}
Each benchmark instance exercises the complete production write path:

\begin{enumerate}
  \item \textbf{Ingestion.} The instance's context is split into chunks of at
        most 32{,}000 characters (conversational corpora are packed
        turn-by-turn into chunks of the same size, preserving every turn) and
        streamed to the agent as consecutive chat turns in one session.
  \item \textbf{Session close.} The session is closed, triggering the
        durable-write chain: an episode summarizer, an LLM fact extractor
        that emits typed facts, and full-fidelity chunk embedding.
  \item \textbf{Sleep consolidation.} One or more sleep cycles run offline
        consolidation: fact deduplication, cross-episode linking,
        co-occurrence edge formation, and graph densification.
  \item \textbf{Answering.} Every question is asked in a \emph{fresh} session.
        The agent answers from its memory (retrieval-augmented context
        injection plus agentic recall tools); the question is framed with the
        benchmark's own per-dataset prompt template.
\end{enumerate}

\subsection{Configuration}
\nous{} ran under its live production configuration (Claude Sonnet~5 as the
reasoning model; hybrid chunk search; typed-edge extraction; cross-encoder
re-ranking; date-aware retrieval; adaptive thinking) with exactly three
eval-hygiene overrides, all disabling background self-improvement features
irrelevant to memory measurement: the heartbeat timer, the scheduler, and
rubric outcome detection. A 117-knob snapshot of the configuration is pinned
in the harness repository for reproducibility.

Two run profiles were used. The \emph{full-fidelity} profile (unbounded
episode summarization, three sleep cycles) is used for moderate contexts.
For multi-megabyte contexts a \emph{big-context} profile bounds the episode
summarizer to 8 chunks, runs one sleep cycle, and paces ingestion turns
(15\,s); this profile eliminated upstream rate-limit stalls and ingest
timeouts entirely (0 errors across all final runs, including 2.3M-character
documents).

\subsection{Isolation}
Every instance gets a fresh \nous{} server process, a unique agent identity,
and a dedicated evaluation database; nothing carries over between instances.
Control probes confirmed the isolation: on Conflict Resolution, the identical
questions asked of an agent \emph{before} ingestion score 0.000, so the
measured accuracy is attributable to memory, not to model priors.

\section{Benchmark and Datasets}

\mab{} streams each dataset's context into the agent and then asks held-out
questions. Table~\ref{tab:datasets} lists the evaluated subsets.

\begin{table}[h]
\centering\small
\begin{tabular}{llrrl}
\toprule
Competency & Source & Context size & Questions used & Grader \\
\midrule
\multirow{4}{*}{CR}
 & factconsolidation\_sh (6k--262k)  & 26k--1.1M chars & $4\times40$ & substring (SQuAD-norm.) \\
 & factconsolidation\_mh (6k--262k)  & 26k--1.1M chars & $4\times40$ & substring (SQuAD-norm.) \\
 & \multicolumn{4}{l}{\footnotesize single-hop (sh) = updated fact; multi-hop (mh) = compose two updated facts} \\[2pt]
\multirow{6}{*}{AR}
 & eventqa\_65536                     & $\sim$285k chars & 200 (5 ctx $\times$ 40) & all-elements recall \\
 & eventqa\_131072                    & $\sim$570k chars & 100 (1 ctx) & all-elements recall \\
 & eventqa\_full                      & $\sim$2.3M chars & 8  & all-elements recall \\
 & SH-Doc QA / MH-Doc QA (=ruler\_qa1/qa2) & 0.9M / 1.9M chars & 100+100 & substring \\
 & longmemeval\_s*                    & 1.6M chars (1282 turns) & 60 (full set, 1 of 5 ctx) & GPT-4o yes/no judge \\[2pt]
\multirow{2}{*}{TTL}
 & icl\_banking77 / clinic150 / nlu   & $\sim$400k chars each & $3\times40$ & parsed-label exact match \\
 & icl\_trec\_coarse / trec\_fine     & $\sim$400k chars each & $2\times40$ & parsed-label exact match \\[2pt]
\multirow{2}{*}{LRU}
 & detective\_qa (10 novels)          & $\sim$400k chars each & 68 & substring \\
 & infbench\_sum\_eng                 & 1.75M chars & 1 & GPT-4o 3-call f1 judge \\
\bottomrule
\end{tabular}
\caption{Evaluated datasets. ``Context size'' is the raw document streamed into
memory. The TTL \emph{recsys} sub-dataset could not be graded faithfully: its
grader requires an entity-id mapping file that is not shipped in the
benchmark's public repository, so TTL here is the in-context-learning family
only.}
\label{tab:datasets}
\end{table}

\section{Testing Technique}

\subsection{Paper-faithful grading}
Each sub-dataset is graded exactly as in the benchmark's \texttt{post\_process}
dispatch:

\begin{itemize}
  \item \textbf{Substring match} (ruler, detective, factconsolidation): the
        SQuAD-style normalization (lowercase, strip punctuation and articles,
        collapse whitespace), then gold-in-prediction containment, max over
        golds.
  \item \textbf{EventQA recall}: \emph{raw} case-insensitive containment of
        every gold element (\texttt{el.lower() in pred.lower()}); the answer is
        correct only if all elements are present. An initial implementation
        that reused the SQuAD normalization here was flagged by independent
        review as strictly more lenient than the paper (punctuation and
        article stripping create false positives) and was corrected before any
        reported run.
  \item \textbf{ICL exact match}: parse the label the prompt requests
        (\texttt{label:~X}), then exact match on the bare label. This is a
        deliberate, conservative interpretation: the benchmark's own ICL
        scoring parses with a different prefix and reports multiple metrics of
        which the substring variant false-positives on digit collisions
        (gold \texttt{5} matches \texttt{label: 15}); our parser cannot.
  \item \textbf{LongMemEval judge}: the benchmark's GPT-4o yes/no judge with
        its verbatim per-question-type prompts (including the
        temporal-reasoning off-by-one tolerance and the abstention protocol
        for \texttt{\_abs} questions).
  \item \textbf{Summarization judge} (InfBench-Sum): the benchmark's three-call
        GPT-4o protocol scoring fluency $f\in\{0,1\}$, key-point recall $r$,
        and sentence-level precision $p$, combined as
        $\mathrm{f1} = f \cdot \frac{2rp}{r+p}$.
\end{itemize}

Answer prompts are likewise the benchmark's own per-dataset templates (e.g.,
the CR template that states the serial-number update rule with few-shot
examples; the eventQA template that frames ``what happened next''). On
identical memory, switching from a generic answer prompt to the benchmark's CR
template moved CR accuracy by +28 percentage points---prompt fidelity is not
optional if numbers are to be compared.

\subsection{Independent review}
All vendored grading and prompting code was audited by a reviewer agent with
fresh context against the benchmark's repository, file by file. The review
confirmed byte-fidelity of the normalizers, judge prompts, dispatch tables,
and the summarization formula; it found exactly one fidelity bug (the eventQA
normalization above), which was fixed and regression-tested before the
reported runs. The harness carries 125 unit tests covering graders, judges,
chunking, persistence, and isolation.

\subsection{Statistical validation}
\label{sec:validation}
Initial runs used 8 questions per source. A dedicated validation pass then
grew the sample until each competency's 95\% CI either cleared the best
published field value or demonstrably did not:

\begin{itemize}
  \item \textbf{TTL corrected downward.} At $n{=}8$/source TTL measured 0.650;
        at $n{=}40$/source ($n{=}200$) it fell to \textbf{0.555}
        (CI [0.486, 0.624]), with one source falling from 0.625 to 0.450.
        The small sample had been an optimistic draw. TTL is therefore
        reported as \emph{at} the field ceiling, not above it.
  \item \textbf{LRU confirmed.} Detective-QA at all 10 novel contexts
        ($n{=}68$) held at \textbf{0.824} versus 0.833 at $n{=}6$; its lower
        CI bound (0.733) clears the best field value (0.622).
  \item \textbf{AR firmed upward.} EventQA-65536 across 5 contexts at
        $n{=}200$ rose to 0.910 versus 0.875 at $n{=}24$, consolidating AR at
        \textbf{0.897} (CI [0.857, 0.936]).
  \item \textbf{CR corrected downward --- via replay, not re-ingest.} At
        $n{=}8$/instance CR measured 0.766; replaying \emph{all 40 questions
        per instance against the same persisted memory} ($n{=}320$; a pure
        sample-size extension with the memory-build variable held fixed,
        content-verified agent$\to$instance mapping) re-priced it to
        \textbf{0.725} (CI [0.676, 0.774]): single-hop 0.887, multi-hop
        0.562, with multi-hop degrading monotonically with context size
        (0.825/0.600/0.525/0.300 at 6k/32k/64k/262k).
  \item \textbf{AR corrected by full coverage --- replay again.} Extending
        every already-ingested AR context to its complete question set
        ($n{=}568$ total) moved every thin cell: LongMemEval fell
        0.875~($n{=}8$)~$\to$~\textbf{0.700}~($n{=}60$, including the first
        measurement of its abstention questions), MH-Doc QA fell
        0.750~$\to$~\textbf{0.710} ($n{=}100$), while eventqa-131072
        \emph{rose} 0.750~$\to$~\textbf{0.860} ($n{=}100$) and SH-Doc QA held
        (0.875~$\to$~0.870, $n{=}100$). Consolidated AR: 0.897~($n{=}232$)
        $\to$ \textbf{0.836} ($n{=}568$, CI [0.806, 0.867]) --- still robustly
        above the benchmark paper's field, no longer clearly above the 2026
        leader.
\end{itemize}

The general lesson---small-sample memory-eval estimates run optimistic; grow
$n$ until the CI clears the comparison band---is recorded as a standing
methodological rule for this project. All results persist per-question to
append-only JSONL as each instance completes, so partial runs are never lost
and every reported number is recomputed from the persisted records.

\section{Results}

\subsection{\nous{} scores}

\begin{table}[h]
\centering
\begin{tabular}{lrrll}
\toprule
Competency & Score & $n$ & 95\% CI & Best published avg.\ / verdict \\
\midrule
Accurate Retrieval        & \textbf{0.836} & 568 & [0.806, 0.867] & 0.718 --- \textbf{above, robust} \\
Conflict Resolution       & \textbf{0.725} & 320 & [0.676, 0.774] & 0.295 --- \textbf{above, robust} \\
Long-Range Understanding  & \textbf{0.824} & 68  & [0.733, 0.914] & 0.622 --- \textbf{above, robust} \\
Test-Time Learning (ICL subset) & \textbf{0.555} & 200 & [0.486, 0.624] & 0.539 --- at ceiling \\
\bottomrule
\end{tabular}
\caption{\nous{} on \mab{} (paper-faithful grading, production configuration,
0 execution errors on all final runs). LRU is detective-QA; the auxiliary
InfBench-Sum summarization judge scored f1 = 0.592 on a single 1.75M-char book
(unfirmed, $n{=}1$). ``Best published avg.'' is the strongest per-competency
average across all methods in the benchmark's results table.
\emph{Comparability tiers:} AR, CR, and LRU are \textbf{T1} (paper's exact
prompt/grader/questions; sole deviation is ingestion transport, Section~
\ref{sec:threats}); TTL is \textbf{T2} (T1 plus a deliberate, conservative
label-parsing choice in the ICL grader). No single aggregate score is
published: TTL omits recsys (grader data file not publicly shipped), and the
giant AR/LRU sources are measured at 1 context each. The repository's
\texttt{BASELINE\_SUMMARY.md} reconciles every number here to its committed
per-question evidence file, tier, and missing-components list.}
\label{tab:results}
\end{table}

Per-source detail: CR ($n{=}320$: the full 40 questions per instance,
replayed against the same persisted memory that produced the initial run)
single-hop 0.887 vs.\ multi-hop 0.562 (the benchmark paper reports at most
$\sim$7\% for multi-hop across its methods); multi-hop degrades with context
size: 0.825/0.600/0.525/0.300 at 6k/32k/64k/262k. AR eventqa
0.910/0.860/0.875 (65k $n{=}200$ / 131k $n{=}100$ / full $n{=}8$), SH-Doc/MH-Doc QA 0.870/0.710 ($n{=}100$ each; MAB's official names --- our source ids ruler\_qa1\_197K/qa2\_421K), longmemeval
0.700 ($n{=}60$); TTL banking77 0.525, clinic150 0.500, nlu 0.425, trec-coarse 0.875,
trec-fine 0.450; LRU detective 56/68.

\subsection{Comparison with the benchmark paper's field (2025)}

Table~\ref{tab:field} reproduces the per-competency averages of representative
methods from the benchmark's results table (its Table~3; the fourth competency
is named Selective Forgetting there and Conflict Resolution in earlier
versions---same datasets). These rows date from the benchmark paper; the 2026
field is covered in Section~\ref{sec:field2026}.

\begin{table}[h]
\centering
\begin{tabular}{llrrrr}
\toprule
Method & Class & AR & TTL & LRU & CR/SF \\
\midrule
GPT-4o-mini (full context) & long-context & 49.2 & 48.6 & 46.2 & 25.0 \\
GPT-4.1-mini (full context) & long-context & 71.8 & 46.2 & 49.1 & 20.5 \\
Claude-3.7-Sonnet (full context) & long-context & 59.7 & \textbf{53.9} & \textbf{62.2} & 22.5 \\
BM25 & RAG & 60.5 & 44.5 & 35.6 & 25.5 \\
Text-Embed-3-Large & RAG & 54.6 & 44.3 & 37.3 & 16.0 \\
HippoRAG-v2 & RAG & 65.1 & 35.8 & 36.2 & \textbf{29.5} \\
Mem0 & agentic memory & 32.6 & 21.2 & 20.7 & 10.0 \\
GraphRAG & structure-aug. & 40.9 & 24.8 & 19.9 & 8.0 \\
MIRIX & agentic memory & 47.5 & 24.1 & 25.4 & 8.0 \\
\midrule
\textbf{\nous{}} (this work) & cognitive memory & \textbf{83.6} & 55.5$^{\S}$ & \textbf{82.4} & \textbf{72.5} \\
\bottomrule
\end{tabular}
\caption{Per-competency averages (\%). Field rows are from the benchmark's
published results; bold marks the best value per column among field methods,
and \nous{} where it exceeds all of them. $^{\S}$~ICL subset only (recsys
excluded). Comparison caveats in
Section~\ref{sec:threats} apply---in particular the base-model mismatch and
the subset sizes.}
\label{tab:field}
\end{table}

Three observations against the paper's table:
\begin{enumerate}
  \item \textbf{Conflict Resolution is the standout vs.\ the 2025 field.}
        \nous{} scores 0.725 where no method in the paper exceeds 0.295, and
        0.562 on multi-hop conflict composition where that field is at most
        $\sim$0.07. We attribute this to \nous{}'s explicit fact store with
        update semantics: conflicting facts are extracted, linked, and
        consolidated during sleep, rather than left as contradictory passages
        for retrieval to stumble over.
  \item \textbf{Retrieval and long-range understanding benefit from
        full-fidelity chunk memory.} AR 0.897 and LRU 0.824 exceed both the
        RAG family and the long-context family, including on 1.6--2.3M-char
        inputs that no context window holds.
  \item \textbf{Test-time learning is the honest gap.} TTL 0.555 edges the
        best value in the paper (0.539) but the CI straddles it. Learning a
        classification skill from thousands of in-context examples is not
        what an episodic/semantic memory is optimized for; accuracy here
        tracks how many labeled exemplars retrieval can surface per query.
\end{enumerate}

\subsection{Comparison with the 2026 field --- sub-dataset level, dual-graded}
\label{sec:field2026}

The field has moved since the benchmark paper. Infini Memory~\cite{infini}
(2026) re-evaluated a modern method suite on the \emph{full} benchmark with
gpt-5-mini as the base model, 4096-token ingestion chunks, and a gpt-5
LLM-as-judge protocol (binary correctness from question, reference, and
output; no judge prompt is published). Every row of its Table~1 was
verbatim-extracted and arithmetically verified (each group average recomputes
from its sub-values; each overall from the four group averages).

To make the comparison protocol-matched in both directions, we additionally
\textbf{re-graded every persisted \nous{} answer with a reconstruction of
their judge protocol} (gpt-5, binary, question + reference + output) --- pure
post-processing over the committed evidence files. Table~\ref{tab:field2026}
reports the field at sub-dataset level with \nous{} under both gradings.

\begin{table}[h]
\centering\scriptsize
\setlength{\tabcolsep}{3.4pt}
\begin{tabular}{lrrrr|r|rr|r|rr|r|rr|r|r}
\toprule
 & \multicolumn{5}{c|}{Accurate Retrieval} & \multicolumn{3}{c|}{TTL} & \multicolumn{3}{c|}{LRU} & \multicolumn{3}{c|}{CR/SF} & \\
Method & SH-QA & MH-QA & LME & Event & Avg & ICL & Rec & Avg & Summ & DetQA & Avg & FC-SH & FC-MH & Avg & Overall \\
\midrule
Mem0        & 28.0 & 36.0 & 42.0 & 35.5 & 35.4 & 35.2 & 11.2 & 23.2 & 12.4 & 42.3 & 27.4 & 25.0 &  3.0 & 14.0 & 25.0 \\
RAPTOR      & 32.0 & 39.0 & 38.2 & 48.2 & 39.4 & 56.8 & 14.0 & 35.4 & 25.3 & 45.6 & 35.5 & 16.0 &  2.0 &  9.0 & 29.8 \\
MemGPT      & 49.0 & 30.0 & 42.7 & 45.6 & 41.8 & 65.6 & 14.3 & 40.0 & 10.9 & 40.2 & 25.6 & 32.0 &  3.0 & 17.5 & 31.2 \\
MemoRAG     & 35.0 & 36.0 & 25.0 & 54.4 & 37.6 & 75.0 & 15.7 & 45.4 & 15.3 & 53.5 & 34.4 & 25.0 &  8.0 & 16.5 & 33.5 \\
LightMem    & 57.0 & 34.0 & 69.2 & 39.2 & 49.9 & 58.0 & 12.7 & 35.4 & 27.5 & 41.0 & 34.2 & 29.0 &  6.0 & 17.5 & 34.2 \\
REMem       & 62.0 & 42.0 & 62.3 & 41.2 & 51.9 & 80.0 & 13.2 & 46.6 & 23.4 & 47.6 & 35.5 & 32.0 &  9.0 & 20.5 & 38.6 \\
HippoRAG-v2 & 78.0 & 70.0 & 54.7 & 72.2 & 68.7 & 65.2 & 12.3 & 38.8 & 32.1 & 54.3 & 43.2 & 58.0 &  5.0 & 31.5 & 45.5 \\
Infini Memory-H & 77.0 & 76.0 & 76.0 & 83.0 & 78.0 & 81.0 & 15.8 & 48.4 & 57.2 & 78.4 & 67.8 & 80.0 & 22.0 & 51.0 & 61.3 \\
Infini Memory-A & 83.0 & 79.0 & 79.3 & 83.6 & 81.2 & \textbf{84.0} & 18.0 & 51.0 & 59.9 & 77.2 & 68.6 & 81.0 & 35.0 & 58.0 & 64.7 \\
\midrule
\nous{} (official graders) & 87.0 & 71.0 & 70.0 & 89.3 & 79.3 & 55.5 & --- & n/a & 59.2 & 82.4 & 70.8 & 88.7 & 56.2 & 72.5 & n/a \\
\nous{} (gpt-5 judge)      & \textbf{91.0} & \textbf{81.0} & 71.7 & \textbf{92.5} & \textbf{84.0} & 62.5 & --- & n/a & 59.2 & \textbf{85.3} & \textbf{72.3} & \textbf{89.4} & \textbf{56.2} & \textbf{72.8} & n/a \\
\bottomrule
\end{tabular}
\caption{2026 field (Infini Memory~\cite{infini} Table~1, full benchmark,
gpt-5-mini base, gpt-5 judge; verbatim-verified) vs.\ \nous{} slices under
both gradings. \nous{} sample sizes: SH-QA/MH-QA $n{=}100$ each (full
question sets), LME $n{=}60$ (full set of its context, incl.\ abstention
questions), Event pooled $n{=}308$, ICL $n{=}200$, Summ $n{=}1$, DetQA
$n{=}68$, FC-SH/FC-MH $n{=}160$ each --- all replayed on the same persisted
memory that produced the original runs. Rec is data-blocked, so \nous{} TTL Avg and
Overall are honestly n/a. \nous{}'s judge is a disclosed reconstruction
(theirs is unpublished). Base models differ (claude-sonnet-5 vs.\
gpt-5-mini).}
\label{tab:field2026}
\end{table}

\paragraph{Backbone-adjusted read.} Every \nous{}-vs-Infini comparison below
carries the backbone confound (claude-sonnet-5 vs.\ gpt-5-mini)$^{\ddagger}$.
The benchmark's own ablations show backbone swings of order $\sim$9.7pp for
agentic-memory methods (MIRIX, GPT-4o-mini$\to$GPT-4.1-mini), so we grade
each margin against that yardstick rather than treating any lead as
architecture-attributable:

\begin{itemize}
  \item \textbf{FC-MH (multi-hop conflict, $n{=}160$): margin $+$21.2pp
        (56.2 vs.\ 35.0) --- exceeds the known backbone swing}; the one
        comparative claim that survives the confound, robust under all three
        grading protocols (0.562 official, 0.562 gpt-5, 0.469 gpt-4o --- all
        above 0.350). It is also the cell Infini's authors identify as unsolved
        (``multi-hop forgetting cannot be settled at write time'').
  \item \textbf{ICL: $-$21.5pp behind (62.5 vs.\ 84.0)} --- the one
        decidable loss; a real capability gap, not a grading artifact,
        plausibly reflecting Infini's up-to-seven-round agentic retrieval
        loop surfacing more labeled exemplars per query.
  \item \textbf{AR: margin $+$2.8pp (84.0 vs.\ 81.2) --- statistical peer.}
        Full coverage collapsed the earlier $+$10.6pp margin: the thin
        $n{=}8$ cells (LME, MH-Doc QA) had been optimistic draws. Within AR,
        Event ($+$8.9) and SH-Doc QA ($+$8.0) lead, MH-Doc QA is a wash
        ($+$2.0), and \textbf{LME is nominally behind ($-$7.6, 71.7 vs.\
        79.3; Wilson 95\% CI on our $n{=}60$ cell [0.58, 0.80])} --- with a
        construct-validity caveat: Infini's table labels LME a
        ``reconstructed multi-session variant,'' which matches \mab{}'s own
        \texttt{longmemeval\_s*} naming but is not confirmed to be the
        identical question set, and our cell covers 1 of 5 LME contexts. We
        therefore report the gap as indicative, not a decided weak cell.
  \item \textbf{FC-SH ($+$8.4pp) and DetQA ($+$8.1pp): within the
        backbone-swing range --- directionally ahead but
        backbone-undecidable}; we do not claim these as architecture wins.
\end{itemize}

Averaging note: the AR ``Avg'' column follows Infini's method (macro mean of
four sub-dataset accuracies) for apples-to-apples --- a method that is
statistically weak on both sides (it weights a 60-item cell equally with a
308-item cell). \nous{}'s item-weighted (micro) AR is 0.836 official / 0.880
judge-graded, vs.\ macro 79.3 / 84.0. Two secondary findings: the
strict-vs-judge delta is $+$0--12.5pp on verbose-answer tasks and
$\approx$0 on short-entity tasks (CR), so official grading was a
conservative floor; and \nous{} multi-hop conflict resolution
\textbf{degrades with context scale} (0.825/0.600/0.525/0.300 at
6k/32k/64k/262k). Whether that degradation is retrieval completeness,
cross-document supersession propagation, or intrinsic multi-hop reasoning
difficulty (the benchmark's own ablation shows o4-mini falling
80\%$\to$14\% on FC-MH as context grows) requires per-failure triage of the
retrieved context; we report the curve as a finding, not a diagnosis.

\section{Threats to Validity}
\label{sec:threats}

\begin{enumerate}
  \item \textbf{Base-model mismatch.} \nous{} reasons with Claude Sonnet~5;
        the benchmark's agentic-memory rows mostly use GPT-4o-mini, and its
        strongest long-context rows use GPT-4.1-mini and Claude-3.7-Sonnet.
        Part of \nous{}'s margin is attributable to the stronger base model.
        The long-context Claude-3.7 row (best field TTL and LRU) is the
        fairest anchor, and \nous{} still exceeds it on AR (+30pp), LRU
        (+20pp), and CR (+54pp).
  \item \textbf{Subsets, not the full benchmark.} Question counts per source
        range from 8 to 200 (Table~\ref{tab:datasets}); the giants
        (eventqa-full, MH-Doc QA 421K, longmemeval) were measured at one context
        each. CIs in Table~\ref{tab:results} reflect this. recsys (TTL) is
        excluded for lack of the grader's data file; InfBench-Sum is a single
        book.
  \item \textbf{Configuration provenance.} The production configuration was
        improved using failures observed on this benchmark's CR subset in an
        earlier phase (write-path fidelity, hybrid chunk retrieval, recall
        depth). The evaluation is therefore best read as ``\nous{} after its
        memory bugs were fixed,'' not as a blind first attempt. No knob was
        tuned per-dataset, and the same configuration serves all four
        competencies.
  \item \textbf{Ingestion transport differs from the paper's protocol.} The
        benchmark feeds context in chunks of 512 tokens for the doc-QA,
        LongMemEval, and conflict-resolution tasks and 4096 tokens elsewhere
        (uniformly 4096 for the commercial memory agents, for cost); our
        harness streams 32{,}000-character turns, and \nous{} re-chunks
        internally for retrieval, so transport granularity is not the
        retrieval granularity. The baseline is therefore paper-faithful in
        prompts, graders, judges, and question sets, but \emph{not} an exact
        reproduction of the paper's ingestion transport. No answer-bearing
        text was lost to this difference: ingest capacity was
        80 $\times$ 32k chars (2.56M), and a post-hoc audit of all 34
        evaluated instances at the exact run settings confirmed
        \textbf{zero dropped chunks} (largest instance: 2.31M chars, 73/80
        chunks; the 1{,}282-turn LongMemEval conversation packed to 50 chunks
        with every turn retained). Runs now stamp
        \texttt{chunks\_sent}/\texttt{chunks\_truncated} on every persisted
        result row, and truncated instances are excluded from headline
        aggregates by construction.
  \item \textbf{Benchmark prompts embed task rules.} The benchmark's own CR
        and eventQA templates tell the agent the update rule / task framing.
        We use them because the field numbers use them; they help any
        system.
  \item \textbf{Judge sharing.} LongMemEval and InfBench-Sum scores depend on
        GPT-4o judges, as in the benchmark.
  \item \textbf{The 2026 comparison is as-reported, not re-run.}
        Table~\ref{tab:field2026} takes the 2026 numbers from the Infini
        Memory paper~\cite{infini} at face value; we did not reproduce them.
        Their protocol differs from the benchmark's official grading (gpt-5
        LLM-judge for all tasks, typically more lenient than substring
        graders; 4096-token chunks) and from ours, and their coverage is the
        full benchmark where ours is slices. We mitigated the grading
        asymmetry by re-grading all persisted \nous{} answers under a
        reconstruction of their judge protocol
        (Table~\ref{tab:field2026}), but their exact judge prompt is
        unpublished and our base model differs, so the comparison is
        indicative, not exact; we make no state-of-the-art claim.
\end{enumerate}

\section{Conclusion}

On the sampled, paper-aligned slices and under the benchmark's exact grading,
\nous{} exceeds every method reported in the \mab{} paper on three of four
memory competencies---Accurate Retrieval (0.897), Conflict Resolution (0.725
at $n{=}320$, a 2.5$\times$ margin over that table's best), and Long-Range
Understanding (0.824)---and matches that table's best Test-Time Learning
(0.555, ICL subset). Against the 2026 field at sub-dataset level, at full
question coverage of every ingested context, under a reconstruction of its
own judge protocol, and with every margin graded against the $\sim$9.7pp
backbone-swing yardstick (Section~\ref{sec:field2026}), the honest placement
is: \textbf{\nous{} and the 2026 leader are statistical peers} --- one
decidable \nous{} win (\textbf{multi-hop conflict resolution}, $+$21.2pp,
tri-protocol robust, on the cell the leader's authors themselves call
unsolved), one decidable loss (\textbf{ICL test-time learning}, $-$21.5pp),
one nominally-behind cell (\textbf{LongMemEval}, $-$7.6pp, comparability
unconfirmed), and every other margin ($+$2.0 to $+$8.9pp) within
backbone-swing range. One dose-response architectural finding stands:
\textbf{multi-hop conflict resolution degrades monotonically with context
size} (0.825/0.600/0.525/0.300 at 6k/32k/64k/262k --- length is the x-axis,
$n{=}40$ per point). The MH-Doc QA and LME corrections, by contrast, are
sampling corrections (their x-axis was $n$, not length; each queries a
single fixed-length context), and we report them as such: our $n{=}8$
estimates were optimistic, full coverage fixed them --- the validation loop
working as intended.
None of this is a state-of-the-art or official-benchmark claim
(Section~\ref{sec:threats}, items 2 and 6). The
conflict-resolution result still suggests that an explicit
extract-link-consolidate memory architecture addresses a failure mode most
retrieval systems share, and the clear TTL gap identifies where \nous{}
should improve next; both justify a full-benchmark evaluation. The harness,
vendored graders, pinned configuration, and per-question result records are
available in the project repository; \texttt{BASELINE\_SUMMARY.md} reconciles
every number to its evidence file and comparability tier.


\appendix
\section{Reconstructed judge prompt (verbatim)}
\label{app:judge}
Infini Memory's judge prompt is unpublished; its protocol is described only as
``a binary correctness judgment based on the question, reference answer, and
model output'' with gpt-5. Our reconstruction, applied verbatim to every
persisted answer (model \texttt{gpt-5}, default temperature, no few-shot):

\begin{quote}\small\ttfamily
You are evaluating whether a model's output correctly answers a question,
given the reference answer.\\[4pt]
Question: \{question\}\\[2pt]
Reference Answer: \{reference\}\\[2pt]
Model Output: \{output\}\\[4pt]
Does the model output contain the correct answer according to the reference?
The output may be verbose or phrased differently; judge semantic correctness
only. Respond with exactly one word: yes or no.
\end{quote}

Multiple reference aliases are joined with `` | ''; outputs are truncated at
4{,}000 characters; correctness = the reply begins with ``yes''. Validation
evidence (judge-vs-official agreement, Cohen's kappa, and a second-judge
sensitivity run with gpt-4o under the identical prompt) is reported in the
repository's \texttt{BASELINE\_SUMMARY.md} alongside the per-question judge
verdict files (\texttt{reports/judge\_regrade/}).

\begin{thebibliography}{9}
\bibitem{mab}
Y.~Hu et al.,
``Evaluating Memory in LLM Agents via Incremental Multi-Turn Interactions''
(\mab{}), ICLR 2026. \texttt{github.com/HUST-AI-HYZ/MemoryAgentBench},
arXiv:2507.05257.
\bibitem{infini}
``Infini Memory: Maintainable Topic Documents for Long-Term LLM Agent
Memory,'' arXiv:2606.10677, 2026.
\end{thebibliography}

\end{document}
